Scattering from compact objects: \\ Regge poles and the Complex Angular Momentum method

### 2019-12-22 rainbowCAM.tex
typo corrected in Eq.A2
removed some material from appendix as it is discussed in 3 c

The remainder of this paper is organised as follows. Sec.~\ref{SecII} reviews the theory of scalar waves on a spherically-symmetric spacetime of a compact object. Here we describe our model spacetime (\ref{SecIIa}), the effective potentials for wave scattering (\ref{subsec:Potentials}) and the physical boundary conditions that define the $S$-matrix (\ref{subsec:bc}). Sec.~\ref{sec:RP} focusses on the Regge pole spectrum. Here we outline the link between Regge poles and QNMs (\ref{subsec:QNMs_RPs}), we describe the numerical method (\ref{subsec:method}), and we present new numerical results for the spectrum (\ref{subsec:results1}) which exhibits three distinct branches of Regge poles (see e.g.~Fig.~\ref{RP_R226_approx_2Mw_3_6_s_1}). In Sec.~\ref{subsec:WKB} we apply the WKB method to obtain an approximate formula for the ``broad resonance'' branch. Sec.~\ref{sec:CAM} concerns the application of CAM techniques to the calculation of the scattering cross section. Here we review the standard partial-wave expansion of the scattering amplitude (\ref{subsec:partial}), we apply CAM techniques to write this as the sum of a residue series and a background integral (\ref{SecIIc}), and we present a selection of numerical results for $d\sigma/d\Omega$ (\ref{subsec:results2}). We conclude with a discussion in Sec.~\ref{sec:conclusions}.

@@

To do this, it is convenient to write the boundary conditions \eqref{bc_1_in} and \eqref{bc_2_in} defining a solution that is regular at the origin in the following form:
\begin{equation}
\label{bc_CAM}
\phi_{\lambda-\frac{1}{2}}(\omega) \sim
\begin{cases}
r^{\lambda+\frac{1}{2}}, &
r \rightarrow 0, \\
\phantom{+} A_{\lambda-1/2}^{(-)}(\omega) e^{-i\omega r_*} & \\ 
 + A_{\lambda-1/2}^{(+)}(\omega) e^{+i\omega r_*} , &r_* \rightarrow +\infty.
\end{cases}
\end{equation}
%
We next consider the radial equation~\eqref{H_Radial_equation} for the regular solution
\begin{equation}\label{Wave_equation_reg}
   \frac{d^2}{dr_*^2}\phi_{\omega,\lambda-1/2}(r)+\left(\omega^2 -V_{\lambda-1/2}\right) \phi_{\omega,\lambda-1/2}(r) =0,
 \end{equation}
and its complex conjugate, 
 \begin{equation}\label{Wave_equation_reg_conjugate}
   \frac{d^2}{dr_*^2}\phi^*_{\omega,\lambda-1/2}(r)+ \left(\omega^2 -V_{\lambda^*-1/2}\right)\phi^*_{\omega,{\lambda-1/2}}(r) = 0 .
 \end{equation}
Multiplying \eqref{Wave_equation_reg} by $\phi^*_{\omega,\lambda-1/2}(r)$ and \eqref{Wave_equation_reg_conjugate} by $\phi_{\omega,\lambda-1/2}(r)$ and taking the difference, it follows that
\begin{widetext}
\begin{equation}\label{Derivative_Wronskian}
\begin{aligned}
 \phi^*_{\omega,\lambda-1/2}(r) \frac{d^2}{dr_*^2}\phi_{\omega,\lambda-1/2}(r) - \phi_{\omega,\lambda-1/2}(r) \frac{d^2}{dr_*^2} \phi^*_{\omega,\lambda-1/2}(r) %&= 2 i|\phi_{\lambda-1/2}(\omega)|^2 f[r]\frac{2\text{Re}[\lambda]\text{Im}[\lambda]}{r^2} \\
                           &= 2 i|\phi_{\lambda-1/2}(\omega)|^2 f[r]\frac{\text{Im}[\lambda^2]}{r^2} .
 \end{aligned}
\end{equation}
\end{widetext}
The LHS of Eq.~\eqref{Derivative_Wronskian} is the derivative of $W$, the Wronskian of the two solutions with respect to $r_*$. The Wronskian is a constant (\textit{i.e.}, $d/dr_* W[\phi^*,\phi] = 0$) if $\lambda$ either real or purely imaginary. According to Eq.~\eqref{Derivative_Wronskian}, we can write
\begin{equation}\label{Derivative_Wronskian_bis}
\frac{d}{dr_*}W[\phi^*,\phi]= 2i|\phi_{\lambda-1/2}(\omega)|^2 f[r]\frac{2\text{Re}[\lambda]\text{Im}[\lambda]}{r^2}.
\end{equation}

@@

Exploring the second and third quadrants, where $\text{Re}[{\lambda}] < 0$, requires a choice to be made on the form of the boundary condition imposed at $r=0$~\eqref{bc_CAM}. A natural choice is to impose regularity, i.e.,~$\phi_{\omega\ell} \sim r^{-\lambda+1/2}$ as $r \rightarrow 0$ for $\text{Re} [\lambda] < 0$. An advantage of this choice is that it leads to the same symmetry as in the black hole case for the $S$-matrix elements (see Ref.~\cite{Andersson:1994rk,Folacci:2019cmc}), viz.,
\begin{equation}\label{Matrix_S_CAM_symm}
S_{-\lambda -1/2}(\omega) =  e^{-2i \pi \lambda} \, S_{\lambda -1/2}(\omega)
\end{equation}
Furthermore, since the wave equation is invariant under the transformation $\lambda \rightarrow -\lambda$ and according to \eqref{Evaluation_W_bis_1} and \eqref{Relation_1st_4th}, it follows that the distribution of poles in the left-half of the complex plane follows by reflection through the origin. In the other words, in the left-half complex $\lambda$-plan (\text{i.e.} $\text{Re}[\lambda] < 0 $), the Regge poles occur for
\begin{equation}
    \begin{dcases}
        \omega>0 \\
        \text{Im}[\lambda]<0
     \end{dcases}
\quad\text{or}\quad
    \begin{dcases}
      \omega<0 \\
        \text{Im}[\lambda]>0
    \end{dcases}
\end{equation}\

>>>>

The remainder of this paper is organised as follows. Sec.~\ref{SecII} reviews the theory of scalar waves on a spherically-symmetric spacetime of a compact object. Here we describe our model spacetime (\ref{SecIIa}), the effective potentials for wave scattering (\ref{subsec:Potentials}) and the physical boundary conditions that define the $S$-matrix (\ref{subsec:bc}). Sec.~\ref{sec:RP} focusses on the Regge pole spectrum. Here we outline the link between Regge poles and QNMs (\ref{subsec:QNMs_RPs}), we describe the numerical method (\ref{subsec:method}), and we present new numerical results for the spectrum (\ref{subsec:results1}) which exhibits three distinct branches of Regge poles (see e.g.~Fig.~\ref{RP_R226_approx_2Mw_3_6_s_1}). In Sec.~\ref{subsec:WKB} we apply the WKB method to obtain an approximate formula for the ``broad resonance'' branch. Sec.~\ref{sec:CAM} concerns the application of CAM techniques to the calculation of the scattering cross section. Here we review the standard partial-wave expansion of the scattering amplitude (\ref{subsec:partial}), we apply CAM techniques to write this as the sum of a residue series and a background integral (\ref{SecIIc}), and we present a selection of numerical results for $d\sigma/d\Omega$ (\ref{subsec:results2}). We conclude with a discussion in Sec.~\ref{sec:conclusions}. In an appendix, we show which quadrants of the complex $\lambda$-plan contain Regge poles of the $S$-matrix.

@@

To do this, it is convenient to write the boundary conditions \eqref{bc_1_in} and \eqref{bc_2_in} defining a solution that is regular at the origin in the following form:
\begin{equation}
\label{bc_CAM}
\phi_{\omega,\lambda-\frac{1}{2}}(r) \sim
\begin{cases}
r^{\lambda+\frac{1}{2}}, &
r \rightarrow 0, \\
\phantom{+} A_{\lambda-1/2}^{(-)}(\omega) e^{-i\omega r_*} & \\
 + A_{\lambda-1/2}^{(+)}(\omega) e^{+i\omega r_*} , &r_* \rightarrow +\infty.
\end{cases}
\end{equation}
%
We next consider the radial equation~\eqref{H_Radial_equation} for the regular solution
\begin{equation}\label{Wave_equation_reg}
   \frac{d^2}{dr_*^2}\phi_{\omega,\lambda-1/2}(r)+\left(\omega^2 -V_{\lambda-1/2}\right) \phi_{\omega,\lambda-1/2}(r) =0,
 \end{equation}
and its complex conjugate,
 \begin{equation}\label{Wave_equation_reg_conjugate}
   \frac{d^2}{dr_*^2}\phi^*_{\omega,\lambda-1/2}(r)+ \left(\omega^2 -V_{\lambda^*-1/2}\right)\phi^*_{\omega,{\lambda-1/2}}(r) = 0 .
 \end{equation}
Multiplying \eqref{Wave_equation_reg} by $\phi^*_{\omega,\lambda-1/2}(r)$ and \eqref{Wave_equation_reg_conjugate} by $\phi_{\omega,\lambda-1/2}(r)$ and taking the difference, it follows that
\begin{widetext}
\begin{equation}\label{Derivative_Wronskian}
\begin{aligned}
 \phi^*_{\omega,\lambda-1/2}(r) \frac{d^2}{dr_*^2}\phi_{\omega,\lambda-1/2}(r) - \phi_{\omega,\lambda-1/2}(r) \frac{d^2}{dr_*^2} \phi^*_{\omega,\lambda-1/2}(r) %&= 2 i|\phi_{\lambda-1/2}(\omega)|^2 f[r]\frac{2\text{Re}[\lambda]\text{Im}[\lambda]}{r^2} \\
                           &= 2 i|\phi_{\lambda-1/2}(\omega)|^2 f[r]\frac{\text{Im}[\lambda^2]}{r^2} .
 \end{aligned}
\end{equation}
\end{widetext}
The LHS of Eq.~\eqref{Derivative_Wronskian} is the derivative of $W$, the Wronskian of the two solutions with respect to $r_*$. The Wronskian is a constant (\textit{i.e.}, $d/dr_* W[\phi^*,\phi] = 0$) if $\lambda$ either real or purely imaginary. According to Eq.~\eqref{Derivative_Wronskian}, we can write
\begin{equation}\label{Derivative_Wronskian_bis}
\frac{d}{dr_*}W[\phi^*,\phi]= 2i|\phi_{\lambda-1/2}(\omega)|^2 f[r]\frac{2\text{Re}[\lambda]\text{Im}[\lambda]}{r^2}.
\end{equation}

@@

%Exploring the second and third quadrants, where $\text{Re}[{\lambda}] < 0$, requires a choice to be made on the form of the boundary condition imposed at $r=0$~\eqref{bc_CAM}. A natural choice is to impose regularity, i.e.,~$\phi_{\omega\ell} \sim r^{-\lambda+1/2}$ as $r \rightarrow 0$ for $\text{Re} [\lambda] < 0$. An advantage of this choice is that it leads to the same symmetry as in the black hole case for the $S$-matrix elements (see Ref.~\cite{Andersson:1994rk,Folacci:2019cmc}), viz.,
%\begin{equation}\label{Matrix_S_CAM_symm}
%S_{-\lambda -1/2}(\omega) =  e^{-2i \pi \lambda} \, S_{\lambda -1/2}(\omega)
%\end{equation}
%Furthermore, since the wave equation is invariant under the transformation $\lambda \rightarrow -\lambda$ and according to \eqref{Evaluation_W_bis_1} and \eqref{Relation_1st_4th}, it follows that the distribution of poles in the left-half of the complex plane follows by reflection through the origin. 
%In the other words, in the left-half complex $\lambda$-plan (\text{i.e.} $\text{Re}[\lambda] < 0 $), the Regge poles occur for

As discussed in Sec.~\ref{subsec:results1}, a good analytic extension of the S-matrix into the left-half complex $\lambda$-plane leads to the symmetry \eqref{Matrix_S_CAM_symm}. Then, it follows that the Regge poles with $\text{Re}[\lambda] < 0 $ may be found for
%
\begin{equation}
    \begin{dcases}
        \omega>0 \\
        \text{Im}[\lambda]<0
     \end{dcases}
\quad\text{or}\quad
    \begin{dcases}
      \omega<0 \\
        \text{Im}[\lambda]>0
    \end{dcases}.
\end{equation}\
